Project Report: \\ Detecting Spam on Social Networking Sites

### 2012-05-08 docs/paper/paper.tex
Latest paper draft.

\section{Introduction}

\subsection{Problem}

Social networks of any significant size see constant spam, scams
and phishing attacks. The nature of these attacks can be quite diverse
and difficult detect. Marketers can spam members with unwanted
advertisements, fraudsters lure users with advance fee frauds and
other confidence tricks, while others attempt to steal user
information by directing users to external phishing pages. To make
matters more difficult, a site with global reach sees communication
among its members in a number foreign languages with varying levels
of ability. This means that much benign content shares characteristics
like misspellings, awkward phrases, etc. that might have made certain
types of common frauds and spam more easy to distinguish on US-based
(or English-language) sites.
% focus on extra impact of non-public features vs. public (bag-of-words)

@@

This project examines in detail the types of malicious and benign
content that are encountered on social networks by analyzing
experimental data available from InterPals, an international social network
for cultural exchange and language practice. For example, the site
attracts a wide variety of financial scams, ranging from Nigerian
"419" scams to romance scams. Another prevalent problem is spam with
links to third-party websites, directing users to various porn/webcam
sites, phishing sites or various untrustworthy online marketplaces.

We will then examine various methods of detecting and preventing abuse
on the site, including those measures that have already been taken
(e.g. various heuristics including IP/location anomaly detection,
frequency capping, duplicate account detection, etc.). However, the
main focus will be on mining experimental data from the site and using
features derived from this data to build and evaluate classifiers to
detect unwanted behavior programmatically. The large volume of data
available to us will provide a unique perspective both on the types of
malicious content that exist on such sites as well as on the
effectiveness of classifier/learning-based approaches to identifying
these activities.

@@

We plan to make heavy use of our unrestricted access to the data of
InterPals, which has over 1.3 million active members. This data
includes a corpus of over a 100 million private messages and another 2 
million messages that have been labeled as spam by users. Other data
includes 40 million or so "wall" comments, 5 million photos, and 8
million photo comments.

\subsection{Techniques}

Using this data, we plan to first identify the most prevalent types of 
malicious activities on the site. We will investigate various machine learning 
techniques to automatically detect these activities, including sampling, clustering and 
classifiers. Sampling and clustering, in addition to the user-labeled and moderator-verified 
spam corpus, will help us identify training data.

Classifiers that we plan to explore 
include:

\begin{itemize}
\item Naive Bayes
\item Regression
\item SVMs 
\end{itemize}

We then test the classifiers on new site data to evaluate their performance.

One important part of this project will be to identify features of undesirable activity that are useful in classification. We plan to examine:

\subsection{Frameworks}
We implement the detection algorithms using Matlab, ScalaNLP, LibSVM, and Spark, an in-memory distributed computing framework designed for machine learning and iterative computation.

@@

The data in our

@@

\subsection{Mixed Features}
\textbf{Account Lifetime.}

@@

% Modified from Reynold's HW1  write-up
We use a na\"ive Bayes model, in which the documents are generated according to the following process.  
First, a class \(c\) is chosen from a probability distribution \(\Pr(c)\).
Then, a document from class \(c\) is generated.
We tried a multinomial model and a Bernoulli model to model this process.
\begin{itemize}
\item \emph{Multinomial model.}  A document length \(m_d\) is chosen and \(m_d\) words \(w_{d,1}, \dotsc, w_{d,m_d}\) are chosen independently from a distribution \(\Pr(w|c)\).
\item \emph{Bernoulli model.}  For each word \(w\) in the vocabulary, \(w\) is added to the document with probability \(\Pr(w|c)\).
\end{itemize}
The distributions \(\Pr(c)\) and \(\Pr(w|c)\) are parameters of the model.

% Modified from Cliff & Antonio's HW2 write-up
Linear regression is a method to model the relationship between an output variable, $y$, and explanatory variables, $X$. The output variable is a linear sum of the explanatory variables multiplied by their corresponding coeffecients, represented by $\beta$. In our case, $y$ represents a review's ranking from one to five, while $X$ is the bag-of-words representation of the review text. This implies that $\beta$ represents the weight of a word's correlation to a higher review. This gives us the simple regression formula: $$y=X\beta$$

In this project, we performed ordinary least squares regression. By minimizing the sum of squared residuals, we can solve for the unkown parameter vector $\beta$ using a closed form solution. $$\beta=(X'X)^{-1}X'y$$ In order to ensure that the matrix $X'X$ is invertible, we perform ridge regression. In ridge regression, we add the identity matrix, $I$, scaled by a factor, $\lambda$, to produce the following equation. $$\beta=(X'X+\lambda I)^{-1}X'y$$ We calculate this $\beta$ for a training set of review text along with their corresponding ratings which range from 1 to 5. We then evaluate our model $\beta$ on the withheld reviews in order to perform a 10-fold cross-validation.

All of our linear regression code was implemented in MATLAB because of its ease-of-use and optimizations on matrix multiplications. We parsed the reviews from the tokenized binary format into a bag-of-words representation stored as a sparse matrix. Matlab uses Compressed Sparse Column format, so we actually stored the matrix $X'$ because the dictionary is very sparse. We also filtered out duplicate reviews by hashing the first ten words in a review to determine distinctness. This filtered the one million reviews to 492,130.

The actual matrix multiplications were done using the sparse $X$ matrix along with our $y$ vector and the ridge parameter. The multiplication $X'X$ takes up too much memory if we use the entire dictionary, so we chose to limit the dictionary size. We found that a dictionary of the 10,000 most frequent words gave sufficient features while also being computable using MATLAB's built-in sparse matrix multiplications. We also tried using a dictionary of 15,000 words, but only had a marginal improvement along with much slower execution. Using our 10,000 word dictionary, we calculated the performance of our matrix multiply to be 0.123 Gflops/s by dividing the number of non-zero elements in our matrices by the time to compute the multiplication.

Since we used ridge regression, we had the ability to tune the $\lambda$ parameter in order to get the best performance. As shown in figure \ref{fig:lambda}, we observed an AUC improvement of about .5\% between the default value of 1 and a tuned parameter of 1500. This is a fairly major difference since the AUC is at 93\% using the default value. Additionally, we also noticed a major difference in the top words when using better $\lambda$ values. Using the default value of 1, our top words included obscure words including ``\_c'', ``\_j'', ``donovan'', and ``matlock.'' When we used the tuned $\lambda$ value of 1500, the words were much more representative, as shown in table \ref{tab:words}.

@@

%intro to svms
SVMs

@@

\begin{tabular}{c|c|c}
    \multicolumn{3}{c}{Unscaled: Data set size, Accuracy (bag-of-words), Accuracy (expert features)}
    Data size & Accuracy, bag-of-words & Accuracy, expert features \\
    \hline
    2k & 0.0 & 0.0 \\
    10k & 0.0 & 0.0 \\
    22k & 0.0 & 0.0 \\ % note we chose 22k because 11k had the recipient info
    50k  & 0.0 & 0.0 \\
    100k  & 0.0 & 0.0 \\
\end{tabular}

\begin{tabular}{c|c|c}
    \multicolumn{3}{c}{Scaled: Data set size, Accuracy (bag-of-words), Accuracy (expert features)}
    Data size & Accuracy, bag-of-words & Accuracy, expert features \\
    \hline
    2k & 0.0 & 0.0 \\
    10k & 0.0 & 0.0 \\
    22k & 0.0 & 0.0 \\ % note we chose 22k because 11k had the recipient info
\end{tabular}

@@

\begin{tabular}{c|c}
    \multicolumn{2}{c}{Tokens most likely to be positive} \\
    token & \(log(p_+ / p_-)\) \\
    \hline
    pos & 0.0 \\
    pos2 & 0.1 \\
\end{tabular}

\begin{tabular}{c|c}
    \multicolumn{2}{c}{Tokens most likely to be negative} \\
    token & \(log(p_+ / p_-)\) \\
    \hline
    neg & 0.0 \\
    neg2 & 0.1 \\
\end{tabular}

@@

Characteristics of spam messages vary widely depending on the type of spam
we currently lump all spam into a single category; we plan to build on the classes we mentioned in section()
We are currently in the process of extending the moderation interface to include more specific tagging capabilities
- certain classifiers might work better on certain types of spam
- clustering to see how well these classes fit the spam types on the site

>>>>

\section{Introduction}
% Modified from project proposal

\subsection{Problem}
Social networks of any significant size see constant spam, scams
and phishing attacks. The nature of these attacks can be quite diverse
and difficult detect. Marketers can spam members with unwanted
advertisements, fraudsters lure users with advance fee frauds and
other confidence tricks, while others attempt to steal user
information by directing users to external phishing pages. To make
matters more difficult, a site with global reach sees communication
among its members in a number foreign languages with varying levels
of ability. This means that much benign content shares characteristics
like misspellings, awkward phrases, etc. that might have made certain
types of common frauds and spam more easy to distinguish on US-based
(or English-language) sites.
% focus on extra impact of non-public features vs. public (bag-of-words)

@@

This paper examines in detail the types of malicious and benign
content that are encountered on social networks by analyzing
experimental data available from InterPals, an international social network
for cultural exchange and language practice. For example, the site
attracts a wide variety of financial scams, ranging from Nigerian
"419" scams to romance scams. Another prevalent problem is spam with
links to third-party websites, directing users to various porn/webcam
sites, phishing sites or various untrustworthy online marketplaces.

We examine various methods of detecting and preventing abuse
on the site, including those measures that have already been taken
(e.g. various heuristics including IP/location anomaly detection,
frequency capping, duplicate account detection, etc.). We then analyze 
message and user account data points to try to differentiate legitimate users
from malicious ones. By mining this data, we extract features 
to build and evaluate classifiers that can detect unwanted behavior 
programmatically. The large volume of data available to us, although we do 
not use all of it, provides a unique perspective both on the types of
malicious content that exist on such sites as well as on the
effectiveness of classifier-based approaches to identifying
these activities.

We investigated several machine learning techniques to automatically 
detect spam and scams in the private message data, and chose to use 
train and evaluate Naive Bayes, Linear Regression, and 
Support Vector Machine (SVM) classifiers. Our implementations used 
a variety of tools, including Matlab, ScalaNLP, LIBSVM, Lucene and Spark, 
an in-memory distributed computing framework designed for machine 
learning and iterative computation.

@@

We enjoyed unrestricted access to the data of InterPals.net, a SNS with 
over 1.3 million active members. This data includes a corpus of over a 
100 million private messages and another 2 million messages that have been 
labeled as spam by users. Other data includes 40 million or so "wall" 
comments, 5 million photos, and 8 million photo comments.

@@

@@

\subsection{Mixed Features}

\textbf{Account Lifetime.}

@@

% Modified from Reynold's HW1  write-up
%We use a na\"ive Bayes model, in which the documents are generated according to the following process.  
%First, a class \(c\) is chosen from a probability distribution \(\Pr(c)\).
%Then, a document from class \(c\) is generated.
%We tried a multinomial model and a Bernoulli model to model this process.
%\begin{itemize}
%\item \emph{Multinomial model.}  A document length \(m_d\) is chosen and \(m_d\) words \(w_{d,1}, \dotsc, w_{d,m_d}\) are chosen independently from a distribution \(\Pr(w|c)\).
%\item \emph{Bernoulli model.}  For each word \(w\) in the vocabulary, \(w\) is added to the document with probability \(\Pr(w|c)\).
%\end{itemize}
%The distributions \(\Pr(c)\) and \(\Pr(w|c)\) are parameters of the model.
%
%
\subsection{Linear Regression}

% Modified from Cliff & Antonio's HW2 write-up
%Linear regression is a method to model the relationship between an output variable, $y$, and explanatory variables, $X$. The output variable is a linear sum of the explanatory variables multiplied by their corresponding coeffecients, represented by $\beta$. In our case, $y$ represents a review's ranking from one to five, while $X$ is the bag-of-words representation of the review text. This implies that $\beta$ represents the weight of a word's correlation to a higher review. This gives us the simple regression formula: $$y=X\beta$$
%
%In this project, we performed ordinary least squares regression. By minimizing the sum of squared residuals, we can solve for the unkown parameter vector $\beta$ using a closed form solution. $$\beta=(X'X)^{-1}X'y$$ In order to ensure that the matrix $X'X$ is invertible, we perform ridge regression. In ridge regression, we add the identity matrix, $I$, scaled by a factor, $\lambda$, to produce the following equation. $$\beta=(X'X+\lambda I)^{-1}X'y$$ We calculate this $\beta$ for a training set of review text along with their corresponding ratings which range from 1 to 5. We then evaluate our model $\beta$ on the withheld reviews in order to perform a 10-fold cross-validation.
%
%All of our linear regression code was implemented in MATLAB because of its ease-of-use and optimizations on matrix multiplications. We parsed the reviews from the tokenized binary format into a bag-of-words representation stored as a sparse matrix. Matlab uses Compressed Sparse Column format, so we actually stored the matrix $X'$ because the dictionary is very sparse. We also filtered out duplicate reviews by hashing the first ten words in a review to determine distinctness. This filtered the one million reviews to 492,130. 
%
%The actual matrix multiplications were done using the sparse $X$ matrix along with our $y$ vector and the ridge parameter. The multiplication $X'X$ takes up too much memory if we use the entire dictionary, so we chose to limit the dictionary size. We found that a dictionary of the 10,000 most frequent words gave sufficient features while also being computable using MATLAB's built-in sparse matrix multiplications. We also tried using a dictionary of 15,000 words, but only had a marginal improvement along with much slower execution. Using our 10,000 word dictionary, we calculated the performance of our matrix multiply to be 0.123 Gflops/s by dividing the number of non-zero elements in our matrices by the time to compute the multiplication. 
%
%
%Since we used ridge regression, we had the ability to tune the $\lambda$ parameter in order to get the best performance. As shown in figure \ref{fig:lambda}, we observed an AUC improvement of about .5\% between the default value of 1 and a tuned parameter of 1500. This is a fairly major difference since the AUC is at 93\% using the default value. Additionally, we also noticed a major difference in the top words when using better $\lambda$ values. Using the default value of 1, our top words included obscure words including ``\_c'', ``\_j'', ``donovan'', and ``matlock.'' When we used the tuned $\lambda$ value of 1500, the words were much more representative, as shown in table \ref{tab:words}.
%

@@

%intro to svms

@@

%\begin{tabular}{c|c|c}
%    \multicolumn{3}{c}{Unscaled: Data set size, Accuracy (bag-of-words), Accuracy (expert features)}
%    Data size & Accuracy, bag-of-words & Accuracy, expert features \\
%    \hline
%    2k & 0.0 & 0.0 \\
%    10k & 0.0 & 0.0 \\
%    22k & 0.0 & 0.0 \\ % note we chose 22k because 11k had the recipient info
%    50k  & 0.0 & 0.0 \\
%    100k  & 0.0 & 0.0 \\
%\end{tabular}
%
%
%\begin{tabular}{c|c|c}
%    \multicolumn{3}{c}{Scaled: Data set size, Accuracy (bag-of-words), Accuracy (expert features)}
%    Data size & Accuracy, bag-of-words & Accuracy, expert features \\
%    \hline
%    2k & 0.0 & 0.0 \\
%    10k & 0.0 & 0.0 \\
%    22k & 0.0 & 0.0 \\ % note we chose 22k because 11k had the recipient info
%\end{tabular}
%

@@

%\begin{tabular}{c|c}
%    \multicolumn{2}{c}{Tokens most likely to be positive} \\
%    token & \(log(p_+ / p_-)\) \\
%    \hline
%    pos & 0.0 \\
%    pos2 & 0.1 \\
%\end{tabular}
%
%\begin{tabular}{c|c}
%    \multicolumn{2}{c}{Tokens most likely to be negative} \\
%    token & \(log(p_+ / p_-)\) \\
%    \hline
%    neg & 0.0 \\
%    neg2 & 0.1 \\
%\end{tabular}
%

@@

Characteristics of spam messages vary widely depending on the type of spam
we currently lump all spam into a single category; we plan to build on the classes we mentioned in section()
We are currently in the process of extending the moderation interface to include more specific tagging capabilities
%- certain classifiers might work better on certain types of spam
%- clustering to see how well these classes fit the spam types on the site
